---
title: 'Reproducible Research: Activity Monitoring Data'
author: "Leonardo Simioni"
date: "2022-09-21"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading and preprocessing the data

#### Loading the data

```{r}
act <- read.csv("activity.csv")
head(act)
```

#### Processing the data into a suitable format for analysis

```{r}
act_nona <- na.omit(act)
head(act_nona)
```

## The mean total number of steps taken per day

#### Calculating and making a histogram of the total number of steps taken each day

```{r}
act_date <- aggregate(steps~date, act_nona, sum)
library(ggplot2)
ggplot(data=act_date, aes(steps)) + geom_histogram(binwidth=500, fill="red", col="black") + labs(title="Total number of steps taken each day", x="total number of steps taken each day")
```

#### Calculating the mean and median of the total number of steps taken per day

```{r}
summary(act_date$steps)
```

## The average daily activity pattern

#### Time series plot of the 5-minute interval and the average number of steps taken

```{r}
act_int <- aggregate(steps~interval, act_nona, mean)
ggplot(data=act_int, aes(x=interval,y=steps)) + geom_line() + labs(title="Average number of steps taken", x="interval (1 interval = 5 minutes)")
```

#### 5-minute interval containing the maximum number of steps

```{r}
act_int[which.max(act_int$steps),]
```

## Imputing missing values

#### Calculating the total number of missing values in the dataset
```{r}
colSums(is.na(act))
```

#### Filling all of the missing values in the dataset

```{r}
act_fill <- act
act_fill$steps <- ifelse(is.na(act_fill$steps) == TRUE, act_int$steps[act_int$interval %in% act_fill$interval], act_fill$steps)
head(act_fill)
```

#### Making a histogram of the total number of steps taken each day and calculating the mean and median total number of steps taken per day
```{r}
act_fill_date <- aggregate(steps~date, act_fill, sum)

p1 <- ggplot(data=act_date, aes(steps)) + geom_histogram(binwidth=500, fill="green", col="black") + labs(x=NULL,y=NULL) + ylim(c(0,12.5))
p2 <- ggplot(data=act_fill_date, aes(steps)) + geom_histogram(binwidth=500, fill="red", col="black") + labs(x=NULL,y=NULL) + ylim(c(0,12.5))

library(ggpubr)
p1_p2 <- ggarrange(p1, p2, labels = c("Without missing data", "With missing data"), ncol = 2, nrow = 1, font.label=list(size=14, face="bold"))
annotate_figure(p1_p2, top = text_grob("Histograms of total number of steps taken each day", face="bold", size=15), bottom = text_grob("total number of steps taken each day"), left = text_grob("count", rot=90))
```

## Differences in activity patterns between weekdays and weekends

#### Creating a vector variable indicating whether a given date is a weekday or weekend day

```{r}
act_week <- act_fill
Sys.setlocale("LC_TIME", "English")

act_week$day <- as.factor(weekdays(as.Date(act_week$date)))
act_week$WD_WE <- as.factor(ifelse(act_week$day == "Saturday" | act_week$day == "Sunday", "weekend", "weekday"))
head(act_week)
```

#### Making a plot containing a time series plot of the 5-minute interval and the average number of steps taken

```{r}
act_week_date <- aggregate(steps ~ interval + WD_WE, act_week, mean)
head(act_week_date)

ggplot(data=act_week_date, aes(colour=WD_WE, x=interval, y=steps)) + geom_line()+labs(x="interval (1 interval = 5 minutes)", y="steps", title="5-minute interval and the average number of steps taken") + facet_grid(WD_WE ~ .)
```


